---
title: "Revenue Forecast - Exploratory Data Analysis Report"
author: "PLZ"
date: "May 12, 2016"
output: html_document
---
0--------1---------2---------3---------4---------5---------6---------7---------8

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
***
### __Objectives:__
> To fully understanding the source data and making appropriate decisions about modeling, need to implement the following objectives:
>
> *	Identify any data attributes with a significant percentage of missing values.
>    +	Determine attributes with missing values be excluded from analysis or imputed
>    +	Recommended logic for imputing values if applicable
>    
> * Identify the data distribution of the initial dataset and any significant outliers related to:
    + Daily Total Amount of Facility Referrals
    + Daily Total Amount of Billed Cases
    + Facility Referrals and/or Billed cases outliers Detection
    + Understand the central tendency and spread of each distribution analysis.
>
> *	Understanding daily trends of the following data attributes:
    +	Facility Referrals
    +	Billed Cases
    +	Summary, Carrier and Facility level trends
>    
> *	Measure the correlation between data attributes in order to find attributes with the greatest influence on Facility Referrals and Billed Cases.
>
> *	Measure elapsed time between Facility Referrals and Billed Cases.
>
> *	Create geography graphs of facility and member locations


***
### __Result:__

#### 1. Missing Value Analysis of Data Attributes
```{r echo=FALSE, Warning=FALSE, message=FALSE}
# Code Part
rm(list=ls())
# Read data extracted from Insight
edadata <- read.csv("C:/Users/pzhou/Analytic Project/RevenueForecast/Data/fullData.csv")
#head(edadata,1)

# Calculate the missing value percentage of each column 
missingvalue.percentage <- data.frame(apply(edadata, 2, function(col)sum(is.na(col))/length(col)*100))
# Add one column "Action"
missingvalue.percentage$Action <- "No Need"
# Rename the column name
colnames(missingvalue.percentage) <- c("MissingValuePercentage","Action")
# Change the action value if missing value percentage exceeds specific value(5%)
missingvalue.percentage$Action[which(missingvalue.percentage$MissingValuePercentage >= 5)] <- "Need to Impute"
missingvalue.percentage$Action[which(missingvalue.percentage$MissingValuePercentage > 0 & missingvalue.percentage$MissingValuePercentage < 5)] <- "Need to Exclude"
# Write the analysis result into csv file
#write.csv(missingvalue.percentage, file = "C:/Users/pzhou/Analytic Project/RevenueForecast/Report/MissingValueAnalysis.csv", row.names = TRUE)
```

*The missing value analysis of each attributes(head 6) shown as below with percentage in descending order.(Here we have set 5% for missing value percentage metrics, that means if percentage exceeds 5%, we need to take some imputation action, but if percentage is less than 5% and greater than zero, consider to take exclusion action, otherwise, no need to take any action.) More detailed, please refer to csv file.*

```{r echo=FALSE, Warning=FALSE, message=FALSE}
# Code Part
# Sort data with missing value percentage in descending order
newdata <- missingvalue.percentage[order(-missingvalue.percentage$MissingValuePercentage),]

# Display the missing value result as table
library(gridExtra)
library(grid)
grid.table(head(newdata), cols = colnames(missingvalue.percentage))
```


#### 2. Daily Facility Referrals Distribution And Outlier Detection

Here is detailed distribution analysis description:

```{r echo=FALSE, Warning=FALSE, message=FALSE}
# Code Part
# Change variable type for Case.Referral.Date and Case.Billing.Month.Date
edadata <- transform(edadata, Case.Referral.Date = as.Date(Case.Referral.Date, format='%m/%d/%Y'), Case.Billing.Month.Date = as.Date(Case.Billing.Month.Date, format='%m/%d/%Y'))

#Calculate the daily facility referral amount
library(plyr)
groupcolumns <- c("Case.Referral.Date")
datacolumns <- c("Case.Referral.Date")
facility.referral.amount <- ddply(edadata, groupcolumns, function(x) count(x[datacolumns]))
#write.csv(facility.referral.amount, file = "C:/Users/pzhou/Analytic Project/RevenueForecast/Report/FacilityReferralAmount.csv", row.names = FALSE)

# Calculate the daily Billed Cases amount
library(plyr)
groupcolumns.cases <- c("Case.Billing.Month.Date")
datacolumns.cases <- c("Case.Billing.Month.Date")
billed.cases.amount <- ddply(edadata, groupcolumns.cases, function(x) count(x[groupcolumns.cases]))
#write.csv(billed.cases.amount, file = "C:/Users/pzhou/Analytic Project/RevenueForecast/Report/BilledCasesAmount.csv", row.names = FALSE)

# Change type to factor
facility.referral.amount.sub <- subset(facility.referral.amount, Case.Referral.Date > as.Date("01/01/2016",format='%m/%d/%Y'))
billed.cases.amount.sub <- subset(billed.cases.amount, Case.Billing.Month.Date > as.Date("01/01/2014",format='%m/%d/%Y'))
facility.referral.amount.sub <- transform(facility.referral.amount.sub, Case.Referral.Date = as.factor(Case.Referral.Date))
billed.cases.amount.sub <- transform(billed.cases.amount.sub, Case.Billing.Month.Date = as.factor(Case.Billing.Month.Date))
```

```{r echo=FALSE, Warning=FALSE, message=FALSE}
# Code Part
# Plot the daily facility referral amount distribution
#plot(facility.referral.amount, xlab = "Time", ylab = "Amount", main = "Daily Facility Referral Amount Distribution")
#hist(facility.referral.amount$freq)

# Plot the density of the daily facility referral amount distribution
#plot(facility.referral.amount, main="Daily Facility Referral Amount Distribution")
#polygon(facility.referral.amount, col="red", border="blue")
  
#Scatterplot and add fit lines
library(car)
scatterplot(freq ~ Case.Referral.Date, data = facility.referral.amount.sub, xlab="Day", ylab="Amount",main="Daily Facility Referral Amount Distribution", labels=row.names(facility.referral.amount.sub))
#abline(lm(facility.referral.amount.sub$freq ~ facility.referral.amount.sub$Case.Referral.Date), col="red") # regression line (y~x) 
lines(lowess(facility.referral.amount.sub$Case.Referral.Date,facility.referral.amount.sub$freq), col="blue") # lowess line (x,y)
```

Here is detailed outlier analysis description

```{r echo=FALSE, Warning=FALSE, message=FALSE}
# Code Part
# Create the time series data
facility.referral.amount.sub.ts <- ts(facility.referral.amount.sub$freq, frequency = 365.25, start = c(2016,1,1))
# Install package tsoutliers
library(tsoutliers)
# Detect outliers with ARIMA model and remove outliers
#facility.referral.amount.sub.ts.outliers.arima <- tso(y = facility.referral.amount.sub.ts, types = c("AO", "LS", "TC"), maxit = 10, remove.method = "bottom-up", tsmethod = "arima", args.tsmethod = list(order = c(5, 1, 0), seasonal = list(order = c(5, 1, 0))))
# Detect the outlier in time series with model linear regression "stsm" and remove outliers
facility.referral.amount.sub.ts.outliers.stsm <- tso(y = facility.referral.amount.sub.ts, types = c("AO", "LS", "TC"), tsmethod = "stsm", args.tsmodel = list(model = "local-level"))
facility.referral.amount.sub.ts.outliers.stsm$fit$call$xreg <- NULL
#facility.referral.amount.sub.ts.outliers.stsm$outliers  # Here we can find that no outliers were detected
# Plot the original series and adjusted series with estimated outlier effect
#plot(facility.referral.amount.sub.ts.outliers.arima, col = "black")  #arima model
plot(facility.referral.amount.sub.ts.outliers.stsm, col = "black")  # linear regression model
```


#### 3. Daily Billed Cases Distribution And Outlier Detection

Here is detailed distribution analysis description:

```{r echo=FALSE, Warning=FALSE, message=FALSE}
# Code Part
# Plot the daily billed cases amount distribution
#plot(billed.cases.amount, xlab = "Time", ylab = "Amount", main = "Daily Billed Cases Amount Distribution")
#hist(billed.cases.amount$freq)

# Plot the density of the daily billed cases amount distribution
plot(billed.cases.amount.sub, main="Daily Billed Cases Amount Distribution")
#polygon(billed.cases.amount, col="red", border="blue")

#Scatterplot
library(car)
scatterplot(freq ~ Case.Billing.Month.Date, data=billed.cases.amount.sub, xlab="Day", ylab="Amount",main="Daily Billed Cases Amount Distribution", labels=row.names(billed.cases.amount.sub))
#abline(lm(billed.cases.amount$freq ~ billed.cases.amount$Case.Billing.Month.Date), col="red") # regression line (y~x) 
#lines(lowess(billed.cases.amount$Case.Billing.Month.Date, billed.cases.amount$freq), col="blue") # lowess line (x,y)
```

Here is detailed outlier analysis description

```{r echo=FALSE, Warning=FALSE, message=FALSE}
# Code Part
# Create the time series data
billed.cases.amount.sub.ts <- ts(billed.cases.amount.sub$freq, frequency = 12, start = c(2014,1))
# Install package tsoutliers
library(tsoutliers)
library(forecast)
# Detect outliers with ARIMA model and remove outliers
#billed.cases.amount.sub.ts.outliers.arima <- tso(y = billed.cases.amount.sub.ts, types = c("AO", "LS", "TC"), maxit = 10, remove.method = "bottom-up", tsmethod = "arima", args.tsmethod = list(order = c(0, 1, 1), seasonal = list(order = c(0, 1, 1))))

# Detect the outlier in time series with model linear regression "stsm" and remove outliers
billed.cases.amount.sub.ts.outliers.stsm <- tso(y = billed.cases.amount.sub.ts, types = c("AO", "LS", "TC"), tsmethod = "stsm", args.tsmodel = list(model = "local-level"))
billed.cases.amount.sub.ts.outliers.stsm$fit$call$xreg <- NULL
billed.cases.amount.sub.ts.outliers.stsm$outliers  # Here we can find that no outliers were detected
# Plot the original series and adjusted series with estimated outlier effect
#plot(billed.cases.amount.ts.outliers.arima, col = "black")  #arima model
plot(billed.cases.amount.sub.ts.outliers.stsm, col = "black", main = "Billed cases")  # linear regression model
```


#### 4. Trend Summary

Here is trend summary 

```{r echo=FALSE, Warning=FALSE, message=FALSE}
# Exclude the missing value of revenue if applicable
#newdata.eld<-na.exclude(newdata) # After exclude, the obs number will decrease 
```

#### 5. Measure the correlation between data attributes and Facility Referrals or Billed Cases

Here is correlation analysis between data attributes

```{r echo=FALSE, Warning=FALSE, message=FALSE}
# Code Part
# Transform all non-numeric to numeric
edadata.new <- transform(edadata, Carrier = as.numeric(Carrier),Case.Assigned.User = as.numeric(Case.Assigned.User),Case.Status = as.numeric(Case.Status),Reporting.Category = as.numeric(Reporting.Category),Actual.Revenue = as.numeric(Actual.Revenue),Facility = as.numeric(Facility),Facility.City = as.numeric(Facility.City),Member.Gender = as.numeric(Member.Gender))
# correlation measure(matrix)
cor(edadata.new[c("Carrier", "Case.Assigned.User", "Case.Status", "Reporting.Category", "Actual.Revenue", "Facility", "Facility.City", "Member.Gender")])

#scatterplot of correlations
#install package psych
library(psych)
pairs.panels(edadata.new[c("Carrier", "Case.Assigned.User", "Case.Status", "Reporting.Category", "Actual.Revenue", "Facility", "Facility.City", "Member.Gender")])
```


#### 6. Measure elapsed time between Facility Referrals and Billed Cases

Here is detailed analysis description

```{r echo=FALSE, Warning=FALSE, message=FALSE}

```

#### 7. Create geography graphs of facility and member locations

Here is detailed analysis description

```{r echo=FALSE, Warning=FALSE, message=FALSE}

```

***
### __Summary and Recommendation:__

Sumany and Recommendation

```{r echo=FALSE, Warning=FALSE, message=FALSE}

```